---
title: "EDA_Italy"
author: "Naylor"
date: "2022-10-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#source('hawkes_functions.R')
library(rnaturalearth)
library(lubridate)
library(raster)
library(sf)
library(sp)
library(ggplot2)
#library(nzsf)
library(ggspatial)
library(rnaturalearthdata)

## This is just the EPSG equivalent of WGS84
crs_wgs84 <- CRS(SRS_string='EPSG:4326')

dataPath <- "../../../data/Italy/"
```

# Get map and extract crs for the map

```{r}
italy.map <- ne_countries(country = 'Italy', returnclass = "sf", scale = 'medium')
italy.crs <- crs(italy.map)

ggplot() + geom_sf(data = italy.map) + theme_bw()

# Get the coastline from the rnaturalearth package for comparison
#italy <- ne_countries(scale = "medium", returnclass = "sf", country = "New Zealand") %>%
#  st_transform(crs = proj_nzsf()) %>%
#  st_crop(get_statistical_areas(area = "EEZ"))
```

# Load and pre-process catalogue

Let's look at the raw catalogue

```{bash Print head of flat file}
head ../../../data/Italy/earlyMedM3.txt
```
Preformat the catalogues...

```{r}

```

Now we can load the catalogue in R and parse the date-times

```{r}
# Load the reformatted catalogue
df_cat <- read.table("../../../data/Italy/earlyMedM3.txt", header=TRUE, sep="|", col.names=c("EventID","DateTime","Latitude","Longitude","DEPTH","Author","Catalog","Contributor","ContributorID","MagType","Magnitude","MagAuthor","EventLocationName","EventType"), fill=TRUE)

df_cat_tmp <- read.table("../../../data/Italy/medLateM3.txt", header=TRUE, sep="|", col.names=c("EventID","DateTime","Latitude","Longitude","DEPTH","Author","Catalog","Contributor","ContributorID","MagType","Magnitude","MagAuthor","EventLocationName","EventType"), fill=TRUE)

df_cat <- rbind(df_cat, df_cat_tmp)

remove(df_cat_tmp)

# add event number
df_cat$event_num <- seq.int(nrow(df_cat))

# parse date-times and set appropriate digital second resoluton
op <- options(digits.secs=7)
options(op)
#df_cat$date <- ymd( paste(df_cat$YR, df_cat$MO, df_cat$DY, sep="-") )
df_cat$dateTime <- ymd_hms( df_cat$DateTime )

# sp version of the catalogue
df_cat.sp <- data.frame(df_cat)
sp::coordinates(df_cat.sp) <- c("Longitude", "Latitude")
crs_wgs84 <- CRS(SRS_string='EPSG:4326')
proj4string(df_cat.sp) <- crs_wgs84

# sf version of the catalogue
df_cat.sf<- st_as_sf(df_cat.sp)

```
```{r}
head(df_cat)
```

# EDA for the whole Mediteranian catalogue

## Histogram of event depths

```{r}
ggplot(df_cat, aes( x="DEPTH" )) + 
  geom_histogram( binwidth=1 ) 
```


```{r Depth histogram}
ggplot(df_cat, aes( x="DEPTH" )) + 
  geom_histogram( binwidth=1 ) + 
  coord_flip (xlim = c(350, 0) ) +
  ggtitle("Hight res. histogram of event depths for depth inversion aretfacts")
```
## Plot of event magnitudes using natural time

```{r}
ggplot(df_cat, aes(x=event_num, y=Magnitude))  + geom_point(size = 0.1) 
```

## Plot of event magnitudes using date-time

```{r Magnitude natural-time series}
ggplot(df_cat, aes(x=dateTime, y=Magnitude)) + 
  geom_point(size = 0.1) +
  ggtitle("Whole Catalogue magnitude time series")
```

```{r Magnitude time series}
# Filtered for M>4
ggplot(df_cat[df_cat$Magnitude>4,], aes(x=dateTime, y=Magnitude)) + 
  geom_point(size = 0.1) +
  ggtitle("Whole catalogue magnitude timeseries for M>4")

```

## Frequency-magnitude analysis for whole Catalogue

```{r National GR Plot}
minMag <- 3
maxMag <- max(df_cat$Magnitude)

mags <- df_cat[df_cat$Magnitude>minMag,]$Magnitude

tmp <- hist(mags, breaks=seq(minMag-0.05,maxMag+0.1,0.1), plot=FALSE)

N.counts <- length( tmp$counts)
tmp$cumulativeCounts <- cumsum(tmp$counts[N.counts:1])[N.counts:1]

m.min <- 4
bin_m.min <- which(tmp$mids==m.min)
freq_m.min <- tmp$counts[bin_m.min]
b <- 1.1
x <- tmp$mids
y <- freq_m.min * 10^(-b*(x-m.min))
y.cum <- tmp$cumulativeCounts[bin_m.min] * 10^(-b*(x-m.min))

ggplot() +
  geom_point( aes(x=tmp$mids, y=tmp$counts) ) +
  geom_point( aes(x=tmp$mids, y=tmp$cumulativeCounts) , color='red', pch="+", size=2) +
  scale_y_log10() +
  ggtitle(paste("Frequency-magnitude plot with arbitary GR dist: b =", b)) +
  xlab("Magnitude") +
  ylab("log10(Frequency)") +
  geom_line(aes(x=x, y=y)) +
  geom_line(aes(x=x, y=y.cum), color='red') +
  geom_vline( xintercept=m.min, lty=2 )

```

```{r National b-value stability}
b.stability.list <- c()
b.error.list <- c()
m.mean <- c()

max.index.x <- length(x)-5

for( i in 1:max.index.x ){
  mag.threshold <- x[i]
  N.events <- length(mags[mags > mag.threshold])
  m.mean <- mean( mags[mags > mag.threshold] )
  m.sd <- sd(mags[mags > mag.threshold])
  b.stability.list[i] <- log10( exp(1) ) / (m.mean - mag.threshold - 0.05)
  b.error.list[i] <- 2.3 * b.stability.list[i]^2 * m.sd / (sqrt(N.events*(N.events-1)))
}

ggplot() +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list) ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list+b.error.list), color=2, lty=2 ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list-b.error.list), color=2, lty=2 ) +
  xlab("Magnitude threshold") +
  ylab("Aki b-value estimate") +
  geom_hline(yintercept = 1, lty=3) +
  ggtitle("b-value stability plot for Whole Catalogue")

```

## Various maps of the events

```{r hexbin event density}
ggplot() +
  geom_hex(data = df_cat[df_cat$Magnitude>3,], aes(x = Longitude, y = Latitude), bins = 50) +
  scale_fill_continuous(type = "viridis") +
  geom_sf(data = italy.map, fill=alpha("lightgrey", 0), color = 'orange', size=0.2) + 
  ggtitle("Density plot for M>3 events") +
  theme_bw()
```


```{r map of locations}
ggplot() +
  geom_sf(data = italy.map, fill=alpha("lightgrey", 0), color = 'green') 
  ggtitle("Map of event locations")
  
```

```{r map of locations}
ggplot() +
  geom_sf(data = df_cat.sf[df_cat$Magnitude>3,], size = 0.05) +
  geom_sf(data = italy.map, fill=alpha("lightgrey", 0), color = 'green') +
  geom_sf(data = df_cat.sf[df_cat$Magnitude>5,], size = 0.5, color='orange') +
  geom_sf(data = df_cat.sf[df_cat$Magnitude>6,], size = 0.5, color='red') +
  ggtitle("Map of event locations")
  
```
# Extract l'Aquila subset 

start.date = "2009-04-06 00:00:00 BST"
end.date = "2010-01-01 00:00:00 BST"
magnitude.completeness = 2.49
min.longitude = 13.00
max.longitude = 13.75
min.latitude = 42.2
max.latitude = 42.5 

```{r Extract l'Aquila subset}
eventDate <- ymd_hms( '2009-04-06 00:00:00 BST' )
endDate <- eventDate + days(400)
startDate <- eventDate - days(50)
deltaLat <- 2.4
latLims <- c( 42.2,42.5)
longLims <- c( 13, 13.75)

minMAG <- 3

# Subset the main catalogue
df_cat.subset <- df_cat[df_cat$Magnitude >= minMAG, ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$Latitude>latLims[1]), ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$Latitude<latLims[2]), ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$Longitude>longLims[1]), ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$Longitude<longLims[2]), ]

head(df_cat.subset)
```

```{r}
ggplot() + 
  geom_point(data=df_cat[df_cat$Magnitude>3,], aes(dateTime, Latitude), size=0.1) +
  geom_point(data=df_cat[df_cat$Magnitude>5,], aes(dateTime, Latitude), size=1.2, color='orange') +
  geom_point(data=df_cat[df_cat$Magnitude>6,], aes(dateTime, Latitude), size=1.5, color='red') +
  ggtitle("L'Aquila latitude-time plot") +
  geom_rect(aes(xmin = as.POSIXct(startDate), xmax = as.POSIXct(endDate), ymin = latLims[1], ymax = latLims[2]), alpha = 0.4, fill='blue', color='blue')
  
```

```{r}
ggplot(df_cat.subset, aes(x=dateTime, y=Magnitude)) + 
  geom_point(size = 0.1) +
  ggtitle("New Zealand magnitude timeseries for M>4") +
  geom_rect( aes(xmin = as.POSIXct(startDate), xmax = as.POSIXct(endDate), ymin = minMAG, ymax = max(df_cat.subset$Magnitude+0.2)), alpha = 0.4, fill=NA, color="blue" )

```


```{r regional GR Plot}
minMag <- 3
maxMag <- max(df_cat.subset$Magnitude, na.rm=TRUE)

mags <- df_cat.subset[df_cat.subset$Magnitude>=minMag,]$Magnitude

tmp <- hist(mags, breaks=seq(minMag-0.05,maxMag+0.4,0.1), plot=FALSE)

N.counts <- length( tmp$counts)
tmp$cumulativeCounts <- cumsum(tmp$counts[N.counts:1])[N.counts:1]

m.min <- 4
bin_m.min <- which(tmp$mids==m.min)
freq_m.min <- tmp$counts[bin_m.min]
b <- 1.1
x <- tmp$mids
y <- freq_m.min * 10^(-b*(x-m.min))
y.cum <- tmp$cumulativeCounts[bin_m.min] * 10^(-b*(x-m.min))

ggplot() +
  geom_point( aes(x=tmp$mids, y=tmp$counts) ) +
  geom_point( aes(x=tmp$mids, y=tmp$cumulativeCounts) , color='red', pch="+") +
  scale_y_log10() +
  ggtitle(paste("Frequency-magnitude plot with arbitary GR dist: b =", b)) +
  xlab("Magnitude") +
  ylab("log10(Frequency)") +
  geom_line(aes(x=x, y=y)) +
  geom_line(aes(x=x, y=y.cum), color='red') +
  geom_vline( xintercept=m.min, lty=2 )

```

```{r}
b.stability.list <- c()
b.error.list <- c()
m.mean <- c()

max.index.x <- length(x)-5

for( i in 1:max.index.x ){
  mag.threshold <- x[i]
  m.mean[i] <- mean( mags[mags > mag.threshold], na.rm=TRUE )
  N.events <- sum(mags > mag.threshold)
  b.stability.list[i] <- log10( exp(1) ) / (m.mean[i] - mag.threshold - 0.05)
  b.error.list[i] <- 2.3 * b.stability.list[i]^2 * sd(mags[mags > mag.threshold]) / (sqrt(N.events*(N.events-1)))
}

ggplot() +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list) ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list+b.error.list), color=2, lty=2 ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list-b.error.list), color=2, lty=2 ) +
  xlab("Magnitude threshold") +
  ylab("Aki b-value estimate") +
  geom_hline(yintercept = 1, lty=3) +
  ggtitle("b-value stability plot for catalogue subset")

```


# Extract l'Amatrice subset 

#######################
## Sequence parameters #
#######################

start.date = "2016-08-24 01:36:00 BST"
end.date = "2017-08-15 19:25:33 BST"
magnitude.completeness = 2.99
min.longitude = 12.9
max.longitude = 13.5
min.latitude = 42.4
max.latitude = 43.08
catalog.path = "data/ISIDE_M2.5.txt"
catalog.header = TRUE
catalog.sep = ' '
catalog.skip = 0
catalog.colnames = NULL


```{r Extract Amatrice subset}
eventDate <- ymd_hms( '2016-08-24 01:36:00 BST' )
endDate <- eventDate + days(400)
startDate <- eventDate - days(50)
deltaLat <- 2.4
latLims <- c( 42.4,43.08)
longLims <- c( 12.9, 13.5)

minMAG <- 3.

# Subset the main catalogue
df_cat.subset <- df_cat[df_cat$Magnitude >= minMAG, ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$Latitude>latLims[1]), ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$Latitude<latLims[2]), ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$Longitude>longLims[1]), ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$Longitude<longLims[2]), ]

head(df_cat.subset)
```

```{r Amatrice latitude-time plot}
ggplot() + 
  geom_point(data=df_cat[df_cat$Magnitude>3,], aes(dateTime, Latitude), size=0.1) +
  geom_point(data=df_cat[df_cat$Magnitude>5,], aes(dateTime, Latitude), size=1.2, color='orange') +
  geom_point(data=df_cat[df_cat$Magnitude>6,], aes(dateTime, Latitude), size=1.5, color='red') +
  ggtitle("Amatrice latitude-time plot") +
  geom_rect(aes(xmin = as.POSIXct(startDate), xmax = as.POSIXct(endDate), ymin = latLims[1], ymax = latLims[2]), alpha = 0.4, fill='blue', color='blue')
  
```

```{r}

ggplot() + 
  geom_point(data=df_cat[df_cat$Magnitude>3,], aes(dateTime, Latitude), size=0.1) +
  geom_point(data=df_cat[df_cat$Magnitude>5,], aes(dateTime, Latitude), size=1.2, color='orange') +
  geom_point(data=df_cat[df_cat$Magnitude>6,], aes(dateTime, Latitude), size=1.5, color='red') +
  ggtitle("New Zealand latitude-time plot") +
  geom_rect(aes(xmin = as.POSIXct(startDate), xmax = as.POSIXct(endDate), ymin = latLims[1], ymax = latLims[2]), alpha = 0.4, fill='blue', color='blue') +
  ylim(latLims)
  
```

```{r}
ggplot(df_cat.subset, aes(x=dateTime, y=Magnitude)) + 
  geom_point(size = 0.1) +
  ggtitle("New Zealand magnitude timeseries for M>4") +
  geom_rect( aes(xmin = as.POSIXct(startDate), xmax = as.POSIXct(endDate), ymin = minMAG, ymax = max(df_cat.subset$Magnitude+0.2)), alpha = 0.4, fill=NA, color="blue" )

```


```{r regional GR Plot}
minMag <- 3
maxMag <- max(df_cat.subset$Magnitude, na.rm=TRUE)

mags <- df_cat.subset[df_cat.subset$Magnitude>=minMag,]$Magnitude

tmp <- hist(mags, breaks=seq(minMag-0.05,maxMag+0.4,0.1), plot=FALSE)

N.counts <- length( tmp$counts)
tmp$cumulativeCounts <- cumsum(tmp$counts[N.counts:1])[N.counts:1]

m.min <- 4
bin_m.min <- which(tmp$mids==m.min)
freq_m.min <- tmp$counts[bin_m.min]
b <- 1.1
x <- tmp$mids
y <- freq_m.min * 10^(-b*(x-m.min))
y.cum <- tmp$cumulativeCounts[bin_m.min] * 10^(-b*(x-m.min))

ggplot() +
  geom_point( aes(x=tmp$mids, y=tmp$counts) ) +
  geom_point( aes(x=tmp$mids, y=tmp$cumulativeCounts) , color='red', pch="+") +
  scale_y_log10() +
  ggtitle(paste("Frequency-magnitude plot with arbitary GR dist: b =", b)) +
  xlab("Magnitude") +
  ylab("log10(Frequency)") +
  geom_line(aes(x=x, y=y)) +
  geom_line(aes(x=x, y=y.cum), color='red') +
  geom_vline( xintercept=m.min, lty=2 )

```

```{r}
b.stability.list <- c()
b.error.list <- c()
m.mean <- c()

max.index.x <- length(x)-5

for( i in 1:max.index.x ){
  mag.threshold <- x[i]
  m.mean[i] <- mean( mags[mags > mag.threshold], na.rm=TRUE )
  N.events <- sum(mags > mag.threshold)
  b.stability.list[i] <- log10( exp(1) ) / (m.mean[i] - mag.threshold - 0.05)
  b.error.list[i] <- 2.3 * b.stability.list[i]^2 * sd(mags[mags > mag.threshold]) / (sqrt(N.events*(N.events-1)))
}

ggplot() +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list) ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list+b.error.list), color=2, lty=2 ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list-b.error.list), color=2, lty=2 ) +
  xlab("Magnitude threshold") +
  ylab("Aki b-value estimate") +
  geom_hline(yintercept = 1, lty=3) +
  ggtitle("b-value stability plot for catalogue subset")

```

