---
title: "EDA_NewZealand"
author: "Naylor"
date: "2022-10-11"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#source('hawkes_functions.R')
library(rnaturalearth)
library(lubridate)
library(raster)
library(sf)
library(sp)
library(ggplot2)

library(nzsf)
library(ggspatial)
library(rnaturalearthdata)

## This is just the EPSG equivalent of WGS84
crs_wgs84 <- CRS(SRS_string='EPSG:4326')

dataPath <- "../../../data/NewZealand/"
```

# Get map and extract crs for the map

```{r Basic map outline}
nz.map <- ne_countries(country = 'New Zealand', returnclass = "sf", scale = 'medium')
nz.crs <- crs(nz.map)

nz_cropped <- st_crop(nz.map, xmin = 130, xmax = 180,
                                    ymin = -70, ymax = -33)
ggplot() + geom_sf(data = nz_cropped) + theme_bw()

# Get the coastline from the rnaturalearth package for comparison
nz <- ne_countries(scale = "medium", returnclass = "sf", country = "New Zealand") %>%
  st_transform(crs = proj_nzsf()) %>%
  st_crop(get_statistical_areas(area = "EEZ"))
```

# Load and pre-process catalogue

Let's look at the raw catalogue

```{bash Print head of flat file}
head ../../../data/NewZealand/hypnzw23R2001_2011.xyze
```

To parse the datetimes, I first pre-process the catalogue to split the MODY and HRMN columns. This makes it easier to parse later on dealine with these as integers etc. Alternatively, I could have loaded as strings.

```{bash preProcess catalogue, eval=FALSE, include=FALSE}
awk -v c=2 '{gsub(/../,"& ",$c); sub(/ $/,"",$c)}1' '../../../data/NewZealand/hypnzw23R2001_2011.xyze' > ../../../data/NewZealand/test.txt 
awk -v c=4 '{gsub(/../,"& ",$c); sub(/ $/,"",$c)}1' '../../../data/NewZealand/test.txt' > ../../../data/NewZealand/test2.txt 
cp ../../../data/NewZealand/test2.txt ../../../data/NewZealand/hypnzw23R2001_2011_MNreformat.xyze
```

Now we can load the catalogue in R and parse the date-times

```{r Cat import and preprocessing}

# Load the reformatted catalogue
df_cat <- read.table("../../../data/NewZealand/hypnzw23R2001_2011_MNreformat.xyze", header=TRUE, sep="", col.names=c("YR","MO","DY","HR","MN","SEC","LATITUDE","LONGITUDE","DEPTH","MAG","NO","RMSRES","x","y","GAP","DMIN","RZDM","NP","NS","SE_OT","SE_H","SE_Z","Q"), fill=TRUE)

# add event number
df_cat$event_num <- seq.int(nrow(df_cat))

# parse date-times and set appropriate digital second resoluton
op <- options(digits.secs=2)
options(op)
df_cat$date <- ymd( paste(df_cat$YR, df_cat$MO, df_cat$DY, sep="-") )
df_cat$dateTime <- ymd_hms( paste(df_cat$YR,"-", df_cat$MO,"-", df_cat$DY," " , df_cat$HR,"-",df_cat$MN,"-" ,df_cat$SEC, sep="") )

# sp version of the catalogue
df_cat.sp <- data.frame(df_cat)
sp::coordinates(df_cat.sp) <- c("LONGITUDE", "LATITUDE")
crs_wgs84 <- CRS(SRS_string='EPSG:4326')
proj4string(df_cat.sp) <- crs_wgs84

# sf version of the catalogue
df_cat.sf<- st_as_sf(df_cat.sp)
```
```{r}
head(df_cat)
```
# EDA for the whole New Zealand catalogue

## Histogram of event depths

```{r Depth histogram}
ggplot(df_cat, aes( x=DEPTH )) + 
  geom_histogram( binwidth=1 ) + 
  coord_flip (xlim = c(350, 0) ) +
  ggtitle("Hight res. histogram of event depths for depth inversion aretfacts")
```
## Plot of event magnitudes using natural time

```{r}
ggplot(df_cat, aes(x=event_num, y=MAG))  + geom_point(size = 0.1) 
```

## Plot of event magnitudes using date-time

```{r Magnitude natural-time series}
ggplot(df_cat, aes(x=dateTime, y=MAG)) + 
  geom_point(size = 0.1) +
  ggtitle("New Zealand magnitude time series")
```

```{r Magnitude time series}
# Filtered for M>4
ggplot(df_cat[df_cat$MAG>4,], aes(x=dateTime, y=MAG)) + 
  geom_point(size = 0.1) +
  ggtitle("New Zealand magnitude timeseries for M>4")

```

## Frequency-magnitude analysis for whole New Zealand

```{r National GR Plot}
minMag <- 1
maxMag <- max(df_cat$MAG)

mags <- df_cat[df_cat$MAG>minMag,]$MAG

tmp <- hist(mags, breaks=seq(minMag-0.05,maxMag+0.1,0.1), plot=FALSE)

N.counts <- length( tmp$counts)
tmp$cumulativeCounts <- cumsum(tmp$counts[N.counts:1])[N.counts:1]

m.min <- 4
bin_m.min <- which(tmp$mids==m.min)
freq_m.min <- tmp$counts[bin_m.min]
b <- 1.1
x <- tmp$mids
y <- freq_m.min * 10^(-b*(x-m.min))
y.cum <- tmp$cumulativeCounts[bin_m.min] * 10^(-b*(x-m.min))

ggplot() +
  geom_point( aes(x=tmp$mids, y=tmp$counts) ) +
  geom_point( aes(x=tmp$mids, y=tmp$cumulativeCounts) , color='red', pch="+", size=2) +
  scale_y_log10() +
  ggtitle(paste("Frequency-magnitude plot with arbitary GR dist: b =", b)) +
  xlab("Magnitude") +
  ylab("log10(Frequency)") +
  geom_line(aes(x=x, y=y)) +
  geom_line(aes(x=x, y=y.cum), color='red') +
  geom_vline( xintercept=m.min, lty=2 )

```

```{r National b-value stability}
b.stability.list <- c()
b.error.list <- c()
m.mean <- c()

max.index.x <- length(x)-5

for( i in 1:max.index.x ){
  mag.threshold <- x[i]
  N.events <- length(mags[mags > mag.threshold])
  m.mean <- mean( mags[mags > mag.threshold] )
  m.sd <- sd(mags[mags > mag.threshold])
  b.stability.list[i] <- log10( exp(1) ) / (m.mean - mag.threshold - 0.05)
  b.error.list[i] <- 2.3 * b.stability.list[i]^2 * m.sd / (sqrt(N.events*(N.events-1)))
}

ggplot() +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list) ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list+b.error.list), color=2, lty=2 ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list-b.error.list), color=2, lty=2 ) +
  xlab("Magnitude threshold") +
  ylab("Aki b-value estimate") +
  geom_hline(yintercept = 1, lty=3) +
  ggtitle("b-value stability plot for New Zealand")

```


## Various maps of the events

```{r hexbin event density}
ggplot() +
  geom_hex(data = df_cat[df_cat$MAG>4,], aes(x = LONGITUDE, y = LATITUDE), bins = 50) +
  scale_fill_continuous(type = "viridis") +
  geom_sf(data = nz_cropped, fill=alpha("lightgrey", 0), color = 'orange', size=0.2) + 
  ggtitle("Density plot for M>4 events") +
  theme_bw()
```




```{r map of locations}
ggplot() +
  geom_sf(data = df_cat.sf[df_cat$MAG>4,], size = 0.05) +
  geom_sf(data = nz_cropped, fill=alpha("lightgrey", 0), color = 'green') +
  geom_sf(data = df_cat.sf[df_cat$MAG>6,], size = 0.5, color='orange') +
  geom_sf(data = df_cat.sf[df_cat$MAG>7,], size = 0.5, color='red') +
  ggtitle("Map of event locations")
  
```

```{r Alt map of locations}
ggplot() +
  geom_sf(data = nz, fill = alpha("lightgrey", 0), colour = "green") +
  geom_sf(data = df_cat.sf[df_cat$MAG>4,], size = 0.05) +
  geom_sf(data = nz, fill = alpha("lightgrey", 0), colour = "green") +
  annotation_north_arrow(location = "tl", which_north = "grid", style = north_arrow_orienteering, height=unit(1., "cm"), width=unit(1., "cm")) +
  annotation_scale(location = "br", unit_category = "metric") +
  geom_sf(data = df_cat.sf[df_cat$MAG>6,], size = 0.5, color='orange') +
  geom_sf(data = df_cat.sf[df_cat$MAG>7,], size = 0.5, color='red', show.legend=points) +
  ggtitle("Map of event locations")
```

## Latitude-time plot for whole New Zealand

```{r Latitude-time plot}
ggplot() + 
  geom_point(data=df_cat[df_cat$MAG>4,], aes(dateTime, LATITUDE), size=0.1) +
  geom_point(data=df_cat[df_cat$MAG>6,], aes(dateTime, LATITUDE), size=1.2, color='orange') +
  geom_point(data=df_cat[df_cat$MAG>7,], aes(dateTime, LATITUDE), size=1.5, color='red') +
  ggtitle("New Zealand latitude-time plot")
```


# Extract the South Island

```{r subset the South Island}
#eventDate <- ymd_hms( '2009-07-15 09:22:29' )

latLims <- c( -42, -47)
minMAG <- 4

# Subset the main catalogue
df_cat.south <- df_cat[df_cat$MAG >= minMAG, ]
df_cat.south <- df_cat.south[ (df_cat.south$LATITUDE<latLims[1]), ]
df_cat.south <- df_cat.south[ (df_cat.south$LATITUDE>latLims[2]), ]
#df_cat.subset <- df_cat.subset[ (df_cat.subset$LONGITUDE>longLims[1]), ]
#df_cat.subset <- df_cat.subset[ (df_cat.subset$LONGITUDE<longLims[2]), ]

endDate <- max(df_cat.south$dateTime)
startDate <- min(df_cat.south$dateTime)
longLims <- c( min(df_cat.south$LONGITUDE), max(df_cat.south$LONGITUDE) )

head(df_cat.south)
```

```{r}
ggplot() +
  geom_sf(data = df_cat.sf[df_cat$MAG>4,], size = 0.05) +
  geom_sf(data = nz_cropped, fill=alpha("lightgrey", 0), color = 'green') +
  geom_sf(data = df_cat.sf[df_cat$MAG>6,], size = 0.5, color='orange') +
  geom_sf(data = df_cat.sf[df_cat$MAG>7,], size = 0.5, color='red') +
  ggtitle("Map of event locations")+
  geom_rect(aes(xmin=longLims[1], xmax=longLims[2], ymin=latLims[1], ymax=latLims[2]), color="blue",fill=NA)
```

```{r}
ggplot() + 
  geom_point(data=df_cat[df_cat$MAG>4,], aes(dateTime, LATITUDE), size=0.1) +
  geom_point(data=df_cat[df_cat$MAG>6,], aes(dateTime, LATITUDE), size=1.2, color='orange') +
  geom_point(data=df_cat[df_cat$MAG>7,], aes(dateTime, LATITUDE), size=1.5, color='red') +
  ggtitle("New Zealand latitude-time plot with Southern Island highlighted") +
  geom_rect(aes(xmin = as.POSIXct(ymd("2001-1-1")), xmax = as.POSIXct(ymd("2012-1-1")), ymin = latLims[1], ymax = latLims[2]), alpha = 0.4, fill='blue')
  
```

```{r}
ggplot(df_cat.south, aes(x=dateTime, y=MAG)) + 
  geom_point(size = 0.1) +
  ggtitle("Southern Island magnitude timeseries for M>4") +
  geom_rect( aes(xmin = as.POSIXct(startDate), xmax = as.POSIXct(endDate), ymin = minMAG, ymax = max(df_cat.south$MAG+0.2)), alpha = 0.4, fill=NA, color="blue" )

```

```{r Southern Island GR Plot}
minMag <- 4
maxMag <- max(df_cat.south$MAG)

mags <- df_cat.south[df_cat.south$MAG>=minMag,]$MAG

tmp <- hist(mags, breaks=seq(minMag-0.05,maxMag+0.1,0.1), plot=FALSE)

N.counts <- length( tmp$counts)
tmp$cumulativeCounts <- cumsum(tmp$counts[N.counts:1])[N.counts:1]

m.min <- 4
bin_m.min <- which(tmp$mids==m.min)
freq_m.min <- tmp$counts[bin_m.min]
b <- 1.1
x <- tmp$mids
y <- freq_m.min * 10^(-b*(x-m.min))
y.cum <- tmp$cumulativeCounts[bin_m.min] * 10^(-b*(x-m.min))

ggplot() +
  geom_point( aes(x=tmp$mids, y=tmp$counts) ) +
  geom_point( aes(x=tmp$mids, y=tmp$cumulativeCounts) , color='red', pch="+") +
  scale_y_log10() +
  ggtitle(paste("Frequency-magnitude plot with arbitary GR dist: b =", b)) +
  xlab("Magnitude") +
  ylab("log10(Frequency)") +
  geom_line(aes(x=x, y=y)) +
  geom_line(aes(x=x, y=y.cum), color='red') +
  geom_vline( xintercept=m.min, lty=2 )

```

```{r Southern Island b-value stability}
b.stability.list <- c()
b.error.list <- c()
m.mean <- c()

max.index.x <- length(x)-5

for( i in 1:max.index.x ){
  mag.threshold <- x[i]
  m.mean[i] <- mean( mags[mags > mag.threshold], na.rm=TRUE )
  N.events <- sum(mags > mag.threshold)
  b.stability.list[i] <- log10( exp(1) ) / (m.mean[i] - mag.threshold - 0.05)
  b.error.list[i] <- 2.3 * b.stability.list[i]^2 * sd(mags[mags > mag.threshold]) / (sqrt(N.events*(N.events-1)))
}

ggplot() +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list) ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list+b.error.list), color=2, lty=2 ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list-b.error.list), color=2, lty=2 ) +
  xlab("Magnitude threshold") +
  ylab("Aki b-value estimate") +
  geom_hline(yintercept = 1, lty=3) +
  ggtitle("b-value stability plot for Southern Island")

```

# Extract the Northern Island

```{r subset the Northern Island}
#eventDate <- ymd_hms( '2009-07-15 09:22:29' )

latLims <- c( -34, -41)
minMAG <- 4

# Subset the main catalogue
df_cat.north <- df_cat[df_cat$MAG >= minMAG, ]
df_cat.north <- df_cat.north[ (df_cat.north$LATITUDE>latLims[2]), ]
#df_cat.north <- df_cat.north[ (df_cat.north$LATITUDE<latLims[2]), ]
#df_cat.subset <- df_cat.subset[ (df_cat.subset$LONGITUDE>longLims[1]), ]
#df_cat.subset <- df_cat.subset[ (df_cat.subset$LONGITUDE<longLims[2]), ]

endDate <- max(df_cat.north$dateTime)
startDate <- min(df_cat.north$dateTime)
longLims <- c( min(df_cat.north$LONGITUDE), max(df_cat.north$LONGITUDE) )

head(df_cat.north)
```

```{r}
ggplot() +
  geom_sf(data = df_cat.sf[df_cat$MAG>4,], size = 0.05) +
  geom_sf(data = nz_cropped, fill=alpha("lightgrey", 0), color = 'green') +
  geom_sf(data = df_cat.sf[df_cat$MAG>6,], size = 0.5, color='orange') +
  geom_sf(data = df_cat.sf[df_cat$MAG>7,], size = 0.5, color='red') +
  ggtitle("Map of event locations")+
  geom_rect(aes(xmin=longLims[1], xmax=longLims[2], ymin=latLims[1], ymax=latLims[2]), color="blue",fill=NA)
```

```{r}
ggplot() + 
  geom_point(data=df_cat[df_cat$MAG>4,], aes(dateTime, LATITUDE), size=0.1) +
  geom_point(data=df_cat[df_cat$MAG>6,], aes(dateTime, LATITUDE), size=1.2, color='orange') +
  geom_point(data=df_cat[df_cat$MAG>7,], aes(dateTime, LATITUDE), size=1.5, color='red') +
  ggtitle("New Zealand latitude-time plot with Northern Island highlighted") +
  geom_rect(aes(xmin = as.POSIXct(ymd("2001-1-1")), xmax = as.POSIXct(ymd("2012-1-1")), ymin = latLims[1], ymax = latLims[2]), alpha = 0.4, fill='blue')
  
```

```{r}
ggplot() + 
  geom_point(data=df_cat.north[df_cat.north$MAG>4,], aes(dateTime, LATITUDE), size=0.1) +
  geom_point(data=df_cat.north[df_cat.north$MAG>6,], aes(dateTime, LATITUDE), size=1.2, color='orange') +
  geom_point(data=df_cat.north[df_cat.north$MAG>7,], aes(dateTime, LATITUDE), size=1.5, color='red') +
  ggtitle("New Zealand latitude-time plot with Northern Island highlighted") +
  geom_rect(aes(xmin = as.POSIXct(ymd("2001-1-1")), xmax = as.POSIXct(ymd("2012-1-1")), ymin = latLims[1], ymax = latLims[2]), alpha = 0.4, fill='blue')
  
```


```{r}
ggplot(df_cat.north, aes(x=dateTime, y=MAG)) + 
  geom_point(size = 0.1) +
  ggtitle("Northern Island magnitude timeseries for M>4") +
  geom_rect( aes(xmin = as.POSIXct(startDate), xmax = as.POSIXct(endDate), ymin = minMAG, ymax = max(df_cat.north$MAG+0.2)), alpha = 0.4, fill=NA, color="blue" )

```

```{r Northern Island GR Plot}
minMag <- 4
maxMag <- max(df_cat.north$MAG)

mags <- df_cat.north[df_cat.north$MAG>=minMag,]$MAG

tmp <- hist(mags, breaks=seq(minMag-0.05,maxMag+0.1,0.1), plot=FALSE)

N.counts <- length( tmp$counts)
tmp$cumulativeCounts <- cumsum(tmp$counts[N.counts:1])[N.counts:1]

m.min <- 4
bin_m.min <- which(tmp$mids==m.min)
freq_m.min <- tmp$counts[bin_m.min]
b <- 1.1
x <- tmp$mids
y <- freq_m.min * 10^(-b*(x-m.min))
y.cum <- tmp$cumulativeCounts[bin_m.min] * 10^(-b*(x-m.min))

ggplot() +
  geom_point( aes(x=tmp$mids, y=tmp$counts) ) +
  geom_point( aes(x=tmp$mids, y=tmp$cumulativeCounts) , color='red', pch="+") +
  scale_y_log10() +
  ggtitle(paste("Frequency-magnitude plot with arbitary GR dist: b =", b)) +
  xlab("Magnitude") +
  ylab("log10(Frequency)") +
  geom_line(aes(x=x, y=y)) +
  geom_line(aes(x=x, y=y.cum), color='red') +
  geom_vline( xintercept=m.min, lty=2 )

```

```{r Northern Island b-value stability}
b.stability.list <- c()
b.error.list <- c()
m.mean <- c()

max.index.x <- length(x)-5

for( i in 1:max.index.x ){
  mag.threshold <- x[i]
  m.mean[i] <- mean( mags[mags > mag.threshold], na.rm=TRUE )
  sum(mags > mag.threshold)
  b.stability.list[i] <- log10( exp(1) ) / (m.mean[i] - mag.threshold - 0.05)
  b.error.list[i] <- 2.3 * b.stability.list[i]^2 * sd(mags[mags > mag.threshold]) / (sqrt(N.events*(N.events-1)))
}

ggplot() +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list) ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list+b.error.list), color=2, lty=2 ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list-b.error.list), color=2, lty=2 ) +
  xlab("Magnitude threshold") +
  ylab("Aki b-value estimate") +
  geom_hline(yintercept = 1, lty=3) +
  ggtitle("b-value stability plot for Northern Island")

```


# Extract the 2009 Fiordland Sequence

UTC time	2009-07-15 09:22:29
ISC event	15157724
USGS-ANSS	ComCat
Local date	15 July 2009
Local time	9:22 pm (NZST)
Magnitude	7.8 Mw
Depth	12.0 km (7.5 mi)
Epicentre	45.762째S 166.562째ECoordinates: 45.762째S 166.562째E
Areas affected	New Zealand
Max. intensity	VI (Strong)
Tsunami	2.3 m (7 ft 7 in)
Aftershocks	>100
Casualties	0

```{r}
eventDate <- ymd_hms( '2009-07-15 09:22:29' )
endDate <- eventDate + days(100)
startDate <- eventDate - days(50)
deltaLat <- 2.4
latLims <- c( -45.762-deltaLat, -45.762+deltaLat)
longLims <- c( 166.562-deltaLat, 166.562+deltaLat)

minMAG <- 4

# Subset the main catalogue
df_cat.subset <- df_cat[df_cat$MAG >= minMAG, ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$LATITUDE>latLims[1]), ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$LATITUDE<latLims[2]), ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$LONGITUDE>longLims[1]), ]
df_cat.subset <- df_cat.subset[ (df_cat.subset$LONGITUDE<longLims[2]), ]

head(df_cat.subset)
```

```{r}
ggplot() +
  geom_sf(data = df_cat.sf[df_cat$MAG>4,], size = 0.05) +
  geom_sf(data = nz_cropped, fill=alpha("lightgrey", 0), color = 'green') +
  geom_sf(data = df_cat.sf[df_cat$MAG>6,], size = 0.5, color='orange') +
  geom_sf(data = df_cat.sf[df_cat$MAG>7,], size = 0.5, color='red') +
  ggtitle("Map of event locations")+
  geom_rect(aes(xmin=longLims[1], xmax=longLims[2], ymin=latLims[1], ymax=latLims[2]), color="blue",fill=NA)
```


```{r}
ggplot() + 
  geom_point(data=df_cat[df_cat$MAG>4,], aes(dateTime, LATITUDE), size=0.1) +
  geom_point(data=df_cat[df_cat$MAG>6,], aes(dateTime, LATITUDE), size=1.2, color='orange') +
  geom_point(data=df_cat[df_cat$MAG>7,], aes(dateTime, LATITUDE), size=1.5, color='red') +
  ggtitle("New Zealand latitude-time plot") +
  geom_rect(aes(xmin = as.POSIXct(ymd("2001-1-1")), xmax = as.POSIXct(ymd("2012-1-1")), ymin = latLims[1], ymax = latLims[2]), alpha = 0.4, fill='blue')
  
```

```{r}
ggplot() + 
  geom_point(data=df_cat[df_cat$MAG>4,], aes(dateTime, LATITUDE), size=0.1) +
  geom_point(data=df_cat[df_cat$MAG>6,], aes(dateTime, LATITUDE), size=1.2, color='orange') +
  geom_point(data=df_cat[df_cat$MAG>7,], aes(dateTime, LATITUDE), size=1.5, color='red') +
  ggtitle("New Zealand latitude-time plot") +
  geom_rect(aes(xmin = as.POSIXct(startDate), xmax = as.POSIXct(endDate), ymin = latLims[1], ymax = latLims[2]), alpha = 0.4, fill='blue')
  
```

```{r}
ggplot(df_cat.subset, aes(x=dateTime, y=MAG)) + 
  geom_point(size = 0.1) +
  ggtitle("New Zealand magnitude timeseries for M>4") +
  geom_rect( aes(xmin = as.POSIXct(startDate), xmax = as.POSIXct(endDate), ymin = minMAG, ymax = max(df_cat.subset$MAG+0.2)), alpha = 0.4, fill=NA, color="blue" )

```

```{r regional GR Plot}
minMag <- 4
maxMag <- max(df_cat.subset$MAG)

mags <- df_cat.subset[df_cat.subset$MAG>=minMag,]$MAG

tmp <- hist(mags, breaks=seq(minMag-0.05,maxMag+0.1,0.1), plot=FALSE)

N.counts <- length( tmp$counts)
tmp$cumulativeCounts <- cumsum(tmp$counts[N.counts:1])[N.counts:1]

m.min <- 4
bin_m.min <- which(tmp$mids==m.min)
freq_m.min <- tmp$counts[bin_m.min]
b <- 1.1
x <- tmp$mids
y <- freq_m.min * 10^(-b*(x-m.min))
y.cum <- tmp$cumulativeCounts[bin_m.min] * 10^(-b*(x-m.min))

ggplot() +
  geom_point( aes(x=tmp$mids, y=tmp$counts) ) +
  geom_point( aes(x=tmp$mids, y=tmp$cumulativeCounts) , color='red', pch="+") +
  scale_y_log10() +
  ggtitle(paste("Frequency-magnitude plot with arbitary GR dist: b =", b)) +
  xlab("Magnitude") +
  ylab("log10(Frequency)") +
  geom_line(aes(x=x, y=y)) +
  geom_line(aes(x=x, y=y.cum), color='red') +
  geom_vline( xintercept=m.min, lty=2 )

```
```{r regional b-value stability}
b.stability.list <- c()
b.error.list <- c()
m.mean <- c()

max.index.x <- length(x)-5

for( i in 1:max.index.x ){
  mag.threshold <- x[i]
  m.mean[i] <- mean( mags[mags > mag.threshold], na.rm=TRUE )
  N.events <- sum(mags > mag.threshold)
  b.stability.list[i] <- log10( exp(1) ) / (m.mean[i] - mag.threshold - 0.05)
  b.error.list[i] <- 2.3 * b.stability.list[i]^2 * sd(mags[mags > mag.threshold]) / (sqrt(N.events*(N.events-1)))
}

ggplot() +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list) ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list+b.error.list), color=2, lty=2 ) +
  geom_line( aes(x=x[1:max.index.x], y=b.stability.list-b.error.list), color=2, lty=2 ) +
  xlab("Magnitude threshold") +
  ylab("Aki b-value estimate") +
  geom_hline(yintercept = 1, lty=3) +
  ggtitle("b-value stability plot for catalogue subset")

```

